---
title: "Accuracy Metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accuracy Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

After verifying documents, ecoreview can calculate how accurately ecoextract found and filled in your records. This guide explains what the metrics mean and how to use them.

# Accessing the Accuracy Modal

Click the **Accuracy** button (bar chart icon) in the header. The modal shows metrics calculated across all verified documents in the connected database.

Metrics are only meaningful for **verified** documents (those with a `reviewed_at` timestamp). Unverified documents have unknown ground truth and are excluded.

# What Gets Measured

The accuracy calculation compares the **LLM's original output** against the **human-corrected ground truth** for each verified document.

Three things can happen to a record during review:

| Event | Meaning |
|-------|---------|
| Record kept unedited | LLM extracted it correctly |
| Record edited | LLM found it but got one or more fields wrong |
| Record deleted | LLM hallucinated a record that doesn't exist |
| Record added by reviewer | LLM missed a real record |

# Metrics

## Record Detection

**Recall** -- what fraction of real records did the LLM find?

$$\text{Recall} = \frac{\text{LLM records kept or edited}}{\text{LLM records kept or edited} + \text{records added by reviewer}}$$

A low recall means the LLM is missing records that are present in the papers.

**Precision** -- what fraction of LLM records were real?

$$\text{Precision} = \frac{\text{LLM records kept or edited}}{\text{LLM records kept or edited} + \text{records deleted by reviewer}}$$

A low precision means the LLM is hallucinating records that don't exist.

**F1** -- harmonic mean of precision and recall. A single summary of detection performance.

## Field-Level Accuracy

Record detection only tells half the story -- a found record may still have incorrect field values. Field-level metrics measure how accurately individual fields were filled in.

**Field precision** -- across all non-deleted, LLM-extracted records, what fraction of field values are correct (unedited)?

**Major edit rate** -- what fraction of records required at least one edit to a "major" field (fields that uniquely identify a record in your schema, or required fields)?

**Average edits per document** -- how many field corrections were needed per verified document on average.

## Interpreting Results

| Metric | Healthy range | Concern |
|--------|--------------|---------|
| Recall | > 0.90 | LLM missing real records |
| Precision | > 0.90 | LLM hallucinating records |
| Field precision | > 0.85 | Fields frequently wrong |
| Major edit rate | < 0.15 | Core fields often incorrect |
| Avg edits/doc | < 3 | Many corrections needed per paper |

These ranges are rough guides -- the right threshold depends on your schema complexity and domain.

# Programmatic Access

Accuracy metrics are also available via `ecoextract::calculate_accuracy()`:

```{r calculate-accuracy}
accuracy <- ecoextract::calculate_accuracy("ecoextract_records.db")

# Detection metrics
accuracy$detection_recall
accuracy$detection_precision
accuracy$detection_f1

# Field-level metrics
accuracy$field_precision
accuracy$major_edit_rate
accuracy$avg_edits_per_document

# Per-document breakdown
accuracy$by_document
```

For a detailed explanation of how each metric is derived, see [ACCURACY.md](https://github.com/n8layman/ecoextract/blob/main/ACCURACY.md) in the ecoextract repository.

# Using Accuracy to Improve Extraction

Accuracy metrics guide prompt and schema iteration:

**Low recall** → the LLM is missing records. Review the extraction prompt -- add examples, clarify what counts as a record, check if papers with many records have lower recall than those with few.

**Low precision** → the LLM is hallucinating. Tighten the schema's required fields and add negative examples to the prompt ("do not extract X if Y is absent").

**Low field precision on specific columns** → those fields are ambiguous in the prompt or schema. Add a definition, examples, and edge-case guidance for those fields.

**High major edit rate** → the unique/required fields are frequently wrong. These are the most impactful errors -- prioritise fixing them in the prompt.

The typical iteration loop:

1. Process a small batch (10-20 papers)
2. Review in ecoreview, click Verify for each
3. Check the Accuracy modal
4. Update `ecoextract/extraction_prompt.md` or `ecoextract/schema.json`
5. Force re-extract the same batch and compare metrics
6. When satisfied, process the full corpus

# Exporting for Analysis

Use **Export Records CSV** to get a flat file suitable for further analysis in R or Excel. The export includes:

- All non-deleted records joined with document metadata
- `record_id` for tracing back to the database
- `reviewed_at` on each row (from the document join) to filter to verified records

```{r export-analysis}
library(readr)
library(dplyr)

records <- read_csv("myproject_records_20260706.csv")

# Verified documents only
verified <- records |> filter(!is.na(reviewed_at))

# Quick summary
cat("Documents verified:", n_distinct(verified$document_id), "\n")
cat("Total records:", nrow(verified), "\n")
```
